draft 1

### 2015-08-05 final write up/draft.rmd
motivation for research writing

---
title: "draft 1"
author: "Amy Cook"
date: "Tuesday, July 07, 2015"
output: word_document
---

@@

Solutions:

Little research has been done into surveying the pricing methods of consulting projects, however

- current practise: literature available on IT and construction projects
        - detailed analysis of project
        - comparison to previous projects by experts - case based reasoning
        - intuition
        - some predictive statistics based on past data 
        - reviewing - trying to find places where things keep going wrong

@@

- has any of this research been mobilised into practise, if not, why?
        - reference class forecasting - pioneered in british government large gov infrastructure

@@

- method for fee pricing has remained the same for decades - detailed analysis of project, add up time based on tasks, allow buffer time for error.

>>>>

---
title: "draft 1"
author: "Amy Cook"
date: "Tuesday, July 07, 2015"
output: word_document
bibliography: Library.bib
---

@@

current practise:

Limited research is available on the cost estimation techniques performed by industry for complex projects, however two surveys were performed in 2000: one on construction projects, the other on IT projects. The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practises. They found that the main method for cost estimation was breaking the project into detailed parts and adding up the cost of each item. The next two most popular methods were comparison with similar projects based on documented facts, and comparison with similar projects based on personal experience. These can all be classified as experience based models [@Akintoye2000]. A survey by Moores and Edwards in 1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project. In the same survey, 91% of software companies cited cost estimation as a problem. Although limited recent literature is available on industry cost estimation practises, historically it can be observed that for complex construction and software projects, cost predictions are chiefly made with a mixture of experience, project comparisons, and an analysis of the project details and tasks. It is also acknowledged that cost estimation is a problem and hypothesised that inexperience and lack of time are chief contributors [@Akintoye2000].

theory about problem with current practise:

@Lovallo2003 tackled the psychology behind the high failure rates of executives in predicting costs of projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development. Their theory stems from Kahneman's work on decision making that won him the nobel prize for economics in 2002. The research argues that humans' natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project. A manager sees challenges in a project as something that can be overcome by the team's high skill level, and downplay or ignore the risk of problems that are out of the team's control. It is for this reason that it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee. Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003]. @Lovallo2003 call the practise of analysing a project based on project details and its unique complexities the 'inside view'. Research has shown that if people are forced to make predictions about their skills after being exposed to an 'outside view', their predictions are significantly more accurate. This outside view can be applied to complex projects, and involves ignoring the details of the current project and instead analysing the outcomes of several projects of a similar description. This is called reference class forecasting [@Lovallo2003]. One method @Lovallo2003 recommend is to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost. The correlation for the current project can then be estimated via a statistical model which is used to adjust the forecasted cost made by detailed analysis (inside view) [@Flyvbjerg2011]. The traditional way to think about a project is to focus on the project details. Even though gathering data from similar projects could significantly improve a cost prediction, it is rarely thought of by managers [@Flyvbjerg2011].

The implementation of reference class forecasting began in project management for the first time in 2004, was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, Australia. An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].

![This method was implemented by the British government for a rail project which is still under completion](C:/Users/n9232371/Documents/github/consult/final write up/UKrail.JPG)

The visualisation presents a powerful communication tool for influencing decision makers and improving forecast cost accuracy, however access to credible data to a sufficient number of projects can be a challenge [@Flyvbjerg2011]. Overall, the theory of optimism bias is a credible explanation for the problem of persistant underestimation of project costs. The accompanying solution of reference class forecasting presents a promising solution given enough accurate data.

@@

        - reference class forecasting - pioneered in british government large gov infrastructure

@@
